### 2020-11-20 Thesis-Chapter2-Background.tex
Write intro to final chapter
Clean up glossary

The weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
war, and increased government surveillance set events in motion toward conflict, but it was the 2013 \ac{snowden}
revelations that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless
wiretapping caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that
caused the real public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption
promised a technical solution. Benefiting from advancements in underlying technologies and under pressure from both the
human rights field and from ordinary users demanding improved privacy, U.S. tech companies responded by introducing
default device encryption for \acl{DAR} and \acl{E2EE} for \acl{DIM} \cite{treguer_us_2018}.

@@

The U.S. corporate response to \ac{snowden} should not be overly construed as a morally motivated defense of civil
rights, as their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the
pursuit of basic digital security. Recall encryption and cryptography's critical role in security as described in
\mysec{sec-crypto-basics}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
security is a central conflict in the \ac{EA} debate. When questioned by lawmakers, technical leaders and executives
have repeatedly appealed to the fact that strong cryptography is a necessity for security \cite{schulze_clipper_2017}.

@@

Written in the aftermath of September 11, the \ac{USAPATRIOTAct} broadly expanded U.S. government surveillance powers
in 25 sections of ``Title II: Enhanced Surveillance Procedures.'' The act notably expands \acs{FISA} order scope
(Section 206), duration (207), and use on U.S. citizens (218). Section 215 enables court orders requiring business
records, but the \ac{snowden} leaks revealed that the government had been secretly using a non-standard definition of
``business records'' in order ``to justify requests for domestic telephone metadata delivered in bulk, not
individualized requests'' \cite{landau_making_2013}.

@@

The \ac{USAFREEDOMAct}, introduced by the same U.S. House representative as the \ac{USAPATRIOTAct}, passed the Senate
almost two years to the day after \ac{snowden}'s leaks began. It contains many reforms to \ac{FISA} and the use of
\acp{NSL}. Most notably, it ends the interpretation of section 215 of the \ac{USAPATRIOTAct} that enabled mass
collection of telephone metadata.

@@

Though key recovery is the most discussed \ac{EA} approach, alternative technical means have been proposed.

@@

% FIXME: Separate out into separate, short chapter. Also adds natural place to introduce my more "advanced" DFD syntax.

>>>>

The weakness of data protection laws and absence of encryption protection laws left the door open for a second crypto
war, and increased government surveillance set events in motion toward conflict, but it was the 2013 Snowden revelations
that constituted the crossing of the threshold. Though policies such as the U.S. \ac{NSA}'s warrantless wiretapping
caused a stir, it was the mass data collection under the agency's \ac{PRISM} and related programs that caused the real
public outcry \cite{landau_making_2013}. With the public interest focused on digital privacy, encryption promised a
technical solution. Benefiting from advancements in underlying technologies and under pressure from both the human
rights field and from ordinary users demanding improved privacy, U.S. tech companies responded by introducing default
device encryption for \acl{DAR} and \acl{E2EE} for \acl{DIM} \cite{treguer_us_2018}.

@@

The U.S. corporate response to Snowden should not be overly construed as a morally motivated defense of civil rights, as
their behavior is a matter of several practical factors \cite{treguer_us_2018}, not least of which is the pursuit of
basic digital security. Recall encryption and cryptography's critical role in security as described in
\mysec{sec-crypto-basics}, and that encryption's dual role as enabler of strong privacy and cornerstone of computer
security is a central conflict in the \ac{EA} debate. When questioned by lawmakers, technical leaders and executives
have repeatedly appealed to the fact that strong cryptography is a necessity for security \cite{schulze_clipper_2017}.

@@

Written in the aftermath of September 11, the \ac{USAPATRIOTAct} broadly expanded U.S. government surveillance powers in
25 sections of ``Title II: Enhanced Surveillance Procedures.'' The act notably expands \acs{FISA} order scope (Section
206), duration (207), and use on U.S. citizens (218). Section 215 enables court orders requiring business records, but
the Snowden leaks revealed that the government had been secretly using a non-standard definition of ``business records''
in order ``to justify requests for domestic telephone metadata delivered in bulk, not individualized requests''
\cite{landau_making_2013}.

@@

The \ac{USAFREEDOMAct}, introduced by the same U.S. House representative as the \ac{USAPATRIOTAct}, passed the Senate
almost two years to the day after Snowden's leaks began. It contains many reforms to \ac{FISA} and the use of \acp{NSL}.
Most notably, it ends the interpretation of section 215 of the \ac{USAPATRIOTAct} that enabled mass collection of
telephone metadata.

@@

Though key escrow is the most discussed \ac{EA} approach, alternative technical means have been proposed.

@@

% FIXME: Separate out into separate, short chapter, probably after the wicked problems chapter. Also adds natural place
% to introduce my more "advanced" DFD syntax.

### 2020-11-20 Thesis-Glossary.tex
Begin threat model chapter
Finish spoofing, begin tampering threats
Finish STRIDE threats
Finish risk analysis
Address review feedback
Clean up glossary

% NOTE: Do not end with a period. The package adds this for you.
% IDEA: auto-update glossary.md. This should be the master, because it is more fully specified

@@

\newterm{snowden}{Snowden}{
    Edward Snowden was an CIA contractor who leaked large volumes of data on NSA programs such as PRISM that performed
    mass collection of internet communications for surveillance
}

@@

\newterm{argmap}{argument map}{
    A graphical representation of the logical structures and relationships between statements, premises, and
    conclusions, used to disentangle complex arguments used in wicked problems
}
% End terms

@@

% Start acronyms
\newabbreviation{AECA}{AECA}{Arms Export Control Act}
\newabbreviation{AV}{AV}{anti-virus}
\newabbreviation{AWA}{AWA}{All Writs Act}
\newabbreviation{CALEA}{CALEA}{Communications Assistance for Law Enforcement Act}
\newabbreviation{CESA}{CESA}{Cyberspace Electronic Security Act}
\newabbreviation{CFAA}{CFAA}{Computer Fraud and Abuse Act}
\newabbreviation{CP}{CP}{child pornography}
\newabbreviation{CSAM}{CSAM}{child sexual abuse material}
\newabbreviation{CSAV}{CSAV}{computer supported argument visualization}
\newabbreviation{DE}{DE}{device encryption}
\newabbreviation{DFD}{DFD}{data flow diagram}
\newabbreviation{DIM}{DIM}{data in motion}
\newabbreviation{DAR}{DAR}{data at rest}
\newabbreviation{E2E}{E2E}{end to end}
\newabbreviation{E2EE}{E2EE}{end to end encryption}
\newabbreviation{EA}{EA}{exceptional access}
\newabbreviation{ECPA}{ECPA}{Electronic Communications Privacy Act}
\newabbreviation{ED}{ED}{ephemeral data}
\newabbreviation{FBI}{FBI}{Federal Bureau of Investigation}
\newabbreviation{FISA}{FISA}{Foreign Intelligence Surveillance Act}
\newabbreviation{FISAAA}{FISAAA}{FISA Amendments Act}
\newabbreviation{FISC}{FISC}{Foreign Intelligence Surveillance Court}
\newabbreviation{GCHQ}{GCHQ}{(UK) Government Communications Headquarters}
\newabbreviation{IBIS}{IBIS}{issue-based information system}
\newabbreviation{ITAR}{ITAR}{International Traffic in Arms Regulations}
\newabbreviation{KEK}{KEK}{key encrypting key}
\newabbreviation{LE}{LE}{law enforcement}
\newabbreviation{LEA}{LEA}{law enforcement agency}
\newabbreviation{LEO}{LEO}{law enforcement officer}
\newabbreviation{NED}{NED}{non-ephemeral data}
\newabbreviation{NSA}{NSA}{National Security Agency}
\newabbreviation{NSL}{NSL}{National Security Letter}
\newabbreviation{PAA}{PAA}{Protect America Act}
\newabbreviation{PGP}{PGP}{Pretty Good Privacy}
\newabbreviation{PIN}{PIN}{personal identification number}
\newabbreviation{PKI}{PKI}{public key infrastructure}
\newabbreviation{PW}{PW}{password}
\newabbreviation{SE}{SE}{strong encryption}
\newabbreviation{SSO}{SSO}{single sign-on}
\newabbreviation{TLS}{TLS}{Transport Layer Security}
\newabbreviation{VPN}{VPN}{virtual private network}
% End acronyms

>>>>

% NOTE: Do not end with a period. The package adds this for you.

@@

% \newterm{snowden}{Snowden}{
%     Edward Snowden was an CIA contractor who leaked large volumes of data on NSA programs such as PRISM that performed
%     mass collection of internet communications for surveillance
% }

@@

\newterm{argmap}{argument map}{
    A graphical representation of the logical structures and relationships between statements, premises, and
    conclusions, used to disentangle complex arguments used in wicked problems
}

\newterm{LDAWMSR}{LDAWMSR}{
    A reference to computer scientist and security researcher Stefan Savage's device key escrow proposal, ``Lawful
    Device Access without Mass Surveillance Risk: A Technical Design Discussion'' \cite{savage_lawful_2018}
}
% End terms

\newcommand{\ldawmsR}[0]{\ii{\ac{LDAWMSR}}}
\newcommand{\ldawmsr}[0]{\ldawmsR~}

@@

% Start acronyms
\newabbreviation{AECA}{AECA}{Arms Export Control Act}
\newabbreviation{APT}{APT}{advanced persistent threat}
\newabbreviation{AV}{AV}{anti-virus}
\newabbreviation{AWA}{AWA}{All Writs Act}
\newabbreviation{CALEA}{CALEA}{Communications Assistance for Law Enforcement Act}
\newabbreviation{CESA}{CESA}{Cyberspace Electronic Security Act}
\newabbreviation{CFAA}{CFAA}{Computer Fraud and Abuse Act}
\newabbreviation{CIA}{CIA}{Central Intelligence Agency}
\newabbreviation{CP}{CP}{child pornography}
\newabbreviation{CSAM}{CSAM}{child sexual abuse material}
\newabbreviation{CSAV}{CSAV}{computer supported argument visualization}
\newabbreviation{DAR}{DAR}{data at rest}
\newabbreviation{DE}{DE}{device encryption}
\newabbreviation{DFD}{DFD}{data flow diagram}
\newabbreviation{DID}{DID}{Device ID}
\newabbreviation{DIM}{DIM}{data in motion}
\newabbreviation{DOS}{DOS}{denial of service}
\newabbreviation{DSK}{DSK}{Device Seal Key}
\newabbreviation{E2E}{E2E}{end to end}
\newabbreviation{E2EE}{E2EE}{end to end encryption}
\newabbreviation{EA}{EA}{exceptional access}
\newabbreviation{ECPA}{ECPA}{Electronic Communications Privacy Act}
\newabbreviation{ED}{ED}{ephemeral data}
\newabbreviation{FBI}{FBI}{Federal Bureau of Investigation}
\newabbreviation{FISA}{FISA}{Foreign Intelligence Surveillance Act}
\newabbreviation{FISAAA}{FISAAA}{FISA Amendments Act}
\newabbreviation{FISC}{FISC}{Foreign Intelligence Surveillance Court}
\newabbreviation{GCHQ}{GCHQ}{Government Communications Headquarters (UK)}
\newabbreviation{HSM}{HSM}{Hardware Security Module}
\newabbreviation{IBIS}{IBIS}{issue-based information system}
\newabbreviation{ITAR}{ITAR}{International Traffic in Arms Regulations}
\newabbreviation{KEK}{KEK}{key encrypting key}
\newabbreviation{LE}{LE}{law enforcement}
\newabbreviation{LEA}{LEA}{law enforcement agency}
\newabbreviation{LEO}{LEO}{law enforcement officer}
\newabbreviation{NED}{NED}{non-ephemeral data}
\newabbreviation{NSA}{NSA}{National Security Agency}
\newabbreviation{NSL}{NSL}{National Security Letter}
\newabbreviation{PAA}{PAA}{Protect America Act}
\newabbreviation{PGP}{PGP}{Pretty Good Privacy}
\newabbreviation{PIN}{PIN}{personal identification number}
\newabbreviation{PKI}{PKI}{public key infrastructure}
\newabbreviation{PW}{PW}{password}
\newabbreviation{SE}{SE}{strong encryption}
\newabbreviation{SSO}{SSO}{single sign-on}
\newabbreviation{TLS}{TLS}{Transport Layer Security}
\newabbreviation{UID}{UID}{Unique ID}
\newabbreviation{VPN}{VPN}{virtual private network}
% End acronyms
